---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
library(here)
```

```{r}
full_data <- read.csv("full_data.csv")
glimpse(full_data)
```

### What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)

```{r}
full_data %>% #no filter needed as NAs already removed
  nrow()
```

### What was the average age of people who are going out trick or treating
```{r}
unique(full_data$age) #checking change to numeric coerced outliers to NA
unique(full_data$going_out) 

full_data %>% 
  filter(going_out == "Yes", 
         age != "Inf",
         age <= 118) %>% #oldest living person
  summarise(avg_age = mean(age))
```
###  What was the average age of people who are not going trick or treating?
```{r}
full_data %>% 
  filter(going_out == "No", 
         age != "Inf",
         age <= 118) %>% 
  summarise(avg_age = mean(age))
```
### For each of joy, despair and meh, which candy bar received the most of these ratings?
```{r}
full_data %>% 
  select(reaction, candy_type) %>% 
  filter(reaction == "JOY") %>% 
  group_by(candy_type) %>% 
  summarise(joy_count=n()) %>% 
  arrange(desc(joy_count)) %>% 
  head(1)

full_data %>% 
  select(reaction, candy_type) %>% 
  filter(reaction == "DESPAIR") %>% 
  group_by(candy_type) %>% 
  summarise(despair_count=n()) %>% 
  arrange(desc(despair_count)) %>% 
  head(1)

full_data %>% 
  select(reaction, candy_type) %>% 
  filter(reaction == "MEH") %>% 
  group_by(candy_type) %>% 
  summarise(meh_count=n()) %>% 
  arrange(desc(meh_count)) %>% 
  head(1)
```
###How many people rated Starburst as despair?
```{r}
full_data %>% 
  filter(candy_type == "Starburst", 
         reaction == "DESPAIR") %>% 
  summarise(star_despair_count = n())
```
###What was the most popular candy bar by this rating system for each gender in the dataset ?
```{r}
full_data_numeric_react <- full_data %>% #setting rating system 
  mutate(reaction = recode(reaction, "JOY" = 1, "MEH" = 0, "DESPAIR" = -1))
```

```{r}
full_data_numeric_react %>% 
  select(reaction, candy_type, gender) %>% 
  filter(gender == "Male") %>% 
  group_by(gender, candy_type) %>% 
  summarise(reaction_total = sum(reaction)) %>% 
  arrange(desc(reaction_total), .by_group = TRUE) %>% 
  head(1)

full_data_numeric_react %>% 
  select(reaction, candy_type, gender) %>% 
  filter(gender == "Female") %>% 
  group_by(gender, candy_type) %>% 
  summarise(reaction_total = sum(reaction)) %>% 
  arrange(desc(reaction_total), .by_group = TRUE) %>% 
  head(1)

full_data_numeric_react %>% 
  select(reaction, candy_type, gender) %>% 
  filter(gender == "Other") %>% 
  group_by(gender, candy_type) %>% 
  summarise(reaction_total = sum(reaction)) %>% 
  arrange(desc(reaction_total), .by_group = TRUE) %>% 
  head(1)

full_data_numeric_react %>% 
  select(reaction, candy_type, gender) %>% 
  filter(gender == "I'd rather not say") %>% 
  group_by(gender, candy_type) %>% 
  summarise(reaction_total = sum(reaction)) %>% 
  arrange(desc(reaction_total), .by_group = TRUE) %>% 
  head(1)
```

###What was the most popular candy bar in each year?
```{r}
full_data_numeric_react %>% 
  select(reaction, candy_type, year) %>% 
  filter(year == "2015") %>% 
  group_by(year, candy_type) %>% 
  summarise(reaction_total = sum(reaction)) %>% 
  arrange(desc(reaction_total), .by_group = TRUE) %>% 
  head(1)

full_data_numeric_react %>% 
  select(reaction, candy_type, year) %>% 
  filter(year == "2016") %>% 
  group_by(year, candy_type) %>% 
  summarise(reaction_total = sum(reaction)) %>% 
  arrange(desc(reaction_total), .by_group = TRUE) %>% 
  head(1)

full_data_numeric_react %>% 
  select(reaction, candy_type, year) %>% 
  filter(year == "2017") %>% 
  group_by(year, candy_type) %>% 
  summarise(reaction_total = sum(reaction)) %>% 
  arrange(desc(reaction_total), .by_group = TRUE) %>% 
  head(1)
```
###What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?
```{r}
full_data_numeric_react %>% 
  select(reaction, candy_type, country) %>% 
  filter(country == "US") %>% 
  group_by(country, candy_type) %>% 
  summarise(reaction_total = sum(reaction)) %>% 
  arrange(desc(reaction_total), .by_group = TRUE) %>% 
  head(1)

full_data_numeric_react %>% 
  select(reaction, candy_type, country) %>% 
  filter(country == "UK") %>% 
  group_by(country, candy_type) %>% 
  summarise(reaction_total = sum(reaction)) %>% 
  arrange(desc(reaction_total), .by_group = TRUE) %>% 
  head(1)

full_data_numeric_react %>% 
  select(reaction, candy_type, country) %>% 
  filter(country == "Canada") %>% 
  group_by(country, candy_type) %>% 
  summarise(reaction_total = sum(reaction)) %>% 
  arrange(desc(reaction_total), .by_group = TRUE) %>% 
  head(1)

full_data_numeric_react %>% 
  select(reaction, candy_type, country) %>% 
  filter(country != "Canada" & country != "US" & country != "UK") %>% 
  group_by(candy_type) %>% #not grouping by country to allow for all others
  summarise(reaction_total_RoW = sum(reaction)) %>% 
  arrange(desc(reaction_total_RoW), .by_group = TRUE) %>% 
  head(1)
```


